\cite{Survey_DL_BCI_2020}
typical workflow: brain signal collection, signal preprocessing, feature extraction, classification and data analysis
signal collection- various method -EEG signals measure the voltage fluctuation resulting from ionic current within the neurons of the brain.
Since the ionic current generated within the brain is measured at the scalp, obstacles (e.g., skull) greatly decrease the signal quality—
the fidelity of the collected EEG signals, measured as Signal-to-Noise Ratio (SNR), is only approximately 5 percent of that of original brain signals

Preprocessing - denoising and enhancement - signal cleaning(smooting noisy signals or resolving inconsistencies)
signal normalization(normalizing each channel of signal along time axis)
signal enhancement(removing direct current)
signal reduction (reduced representation of the signal)

Feature extraction: extracting discriminating features from the input signals through domain knowledge
usually extracted from time-domain, frequency-domain, and time-frequency domain.
manual feature extraction - time consuming and difficult.

Classification: users psychological or physical status-for further info analysis.

DeepLearning: 1, For feature extraction. 2, for classification. 3,for both extraction and classification

Why deep learning?
Brain waves corrupted by various biological and environmental artifacts. Hence distill inforamtive data from noisy signal.
low SNR of non stationary brain signals. Classical machine learning focuses on static data  hence cannot classify rapidly changing brain signals accurately.

DNNs can capture both high level features and latent dependencies through deep structures.

Brain Imaging techniques


\cite{2004_EEG_ML}
Study on Machine learning methods with particular focus on feature selection. Dicusses various flaws in validating the machine learning methods and how classification is linked to the task of robust feature selection.
BCI commonly uses linear classifiers, but can go wrong if the underlying assumptions are not valid.example: presence of outliers or strong noise.
Regularization is a must.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Another cool thing about \LaTeX~is its referencing system. This template is set up to use harvard-style referencing. You can do this by using \texttt{\textbackslash citep\{citekey\}}. It will print out something like this: \citep{aad2012observation}. Or alternatively, you can use \texttt{\textbackslash cite\{citekey\}} to cite things like this: \cite{chatrchyan2012observation}. This template uses Bib\LaTeX~for referencing, with a Biber backend. This is primarily due to the extensive features Bib\LaTeX~provides, along with the option of glossaries. If you want to customise the referencing style, you can either modify the template slightly to use different options, or use \texttt{\textbackslash usepackage} again to reimport it. There's probably some commands to change its options after its been imported too.

% \section{Ludography}
% This thesis template also contains an optional ludography. This is primarily for Games Development students, who wish to cite games in their thesis. To use this, just put references into your bib file as usual with the game's details. Then, make sure \texttt{keywords} is set to \texttt{\{game\}}. This is what is used to determine which references are games, and which are actual papers. For a more elaborate example, see \texttt{bib/ludography.bib}.

% Also, make sure that the \texttt{title} key is actually the author of the game, and the \texttt{author} is the title of the game. The reason this is swapped around is because Bib\LaTeX~likes to print references out with the author first. Then, just add \texttt{\textbackslash printLudography} with an optional title argument to print out all citations like \texttt{\textbackslash printLudography} or \texttt{\textbackslash printLudography[Games]}.

% You can also use the \texttt{ludography} environment if you wish to print out some text before the list of games is printed. An example of this can be seen in \texttt{main.tex}. To cite games, you can \texttt{\textbackslash cite} it like any other reference. However, if you want it to display the title instead of the standard referencing style, you can use \texttt{\textbackslash citeGame} instead.

% Here is an example of a cited game with a normal reference style: ~\cite{spaceinvaders}. Ugh, pretty ugly. Instead, here the two are  cited in the next sentence as games with \texttt{\textbackslash citeGame}. Both \citeGame{spaceinvaders} and \citeGame{breakout} were games made by Atari. Much better!